---
title: "Introduction to Statistical Learning: Lab Chapter 2"
output: html_notebook
---

---

###Basic Commands

---

Use `c()` to create a vector:

```{r}
x <- c(1, 3, 2, 5)
```

Use `<-` or `=` to save content to an object:

```{r}
x = c(1, 6, 2)

y <- c(1, 4, 3)
```

Use `length()` to check the length of vector:

```{r}
length(x)

length(y)
```

Adding two vectors of similar length will add the elements in the first vector to the respective elements in the second vector:

```{r}
x + y
```

Use `ls()` to look at a list of all objects in the current working environment:

```{r}
ls()
```

Use `rm()` to delete any elements in the working environment:

```{r}
rm(x, y)

ls()
```

Use `rm()` with `ls()` to remove all objects at once:

```{r}
rm(list=ls())
```

Use `matrix()` to create a matrix of numbers:

```{r}
x <- matrix(data = c(1,2,3,4), nrow = 2, ncol = 2)
```

Use `sqrt` to return the square root of each element in a vector or matrix:

```{r}
sqrt(x)
```

Use `rnorm()` to generate a vector of random normal variables:

```{r}
x <- rnorm(n = 10)
y <- x + rnorm(n = 10, mean = 50, sd = 0.1)
```

Use `cor()` to compute the correlation between two vectors:

```{r}
cor(x, y)
```

Use `set.seed()` to set the seed for random number generators.  This ensures code is reproducible:

```{r}
set.seed(4)
```

Use `mean()` , `var()` , and `sd()` to compute the mean, variance and standard deviation of a vector respectively:

```{r}
mean(x)
var(x)
sd(x)
```

---

###Graphics

---

Use the `plot()` function as the basic way to produce plots in R:

```{r}
x <- rnorm(100)
y <- rnorm(100)
plot(x, y)
plot(x, y, xlab = "this is the x-axis", ylab = "this is the y-axis", main = "this is the title")
```

Use `pdf()` or `jpeg()` to save plots to file:

```{r eval = FALSE}
pdf("Figure.pdf")
plot(x, y)
dev.off()
```

The function `dev.off()` indicates to R that we are done with the plot we are saving to file.

Use `seq(a, b)` to create a sequence of numbers between `a` and `b`:

```{r}
x <- seq(1, 10)
x
```


USe `contour()` to create a contour plot representing 3-dimensional data:

```{r}
y <- x

f <- outer(x, y, function(x, y) cos(y) / (1 + x ^ 2))
contour(x, y, f)
```

Use `image()` to create a heatmap in base R:

```{r}
image(x, y, f)
```

Use `persp()` to create a 3-dimensional plot in base R:

```{r}
persp(x, y, f)
```

---

###Indexing Data

---

Given a matrix `A`:

```{r}
A <- matrix(1:16, 4, 4)
```

Use `A[2,3]` to select the element in row 2 column 3 of the matrix:

```{r}
A[2,3]
```

Use `A[1, ]` to select all elements in row 1 of the matrix:

```{r}
A[1, ]
```

Use a `-` sign in the index to tell R to keep all elements except those indicated in the index:

```{r}
A[-1, ]
```

Use `dim()` to output the number of rows and columns of a matrix:

```{r}
dim(A)
```


---

###Loading Data

---

This section contains mostly unused base R commands.  


---

###Additional Graphical and Numerical Summaries

---

Use `pairs()` to craete a scatterplot matrix for every pair of data in a given data set.

```{r}
pairs(mtcars)
```

Use `identify()` to make base R plots more interactive enabing one to print the value of a variable of interest by clicking it on a plot.

Use `summary()` to produce a numerical summary of each variable in a particular data set:

```{r}
summary(mtcars)
```

Use `q()` to quite R.





